---
title: "01_dataset_creation"
author: "Claudia Jiménez"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
 output:
   html_notebook:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes
---
# Environment

```{r message=FALSE, warning=FALSE}
library(bigrquery)
library(DBI)
library(dplyr)
library(summarytools)
library(tidyverse)
```
# BigQuery related functions

This chunks creates the run_query and getSQL function.

```{r}
# Function that takes in a SQL command and runs it on BigQuery. This avoids the connection data in each iteration
run_query<-function(query){
  query_output<-dbGetQuery(con,query)
  return(query_output)
}

# Function for reading SQL files from a folder
getSQL <- function(filepath) {
  con = file(filepath, "r")
  sql.string <- ""
  while (TRUE) {
    line <- readLines(con, n = 1)
    if (length(line) == 0) {
      break
    }
    line <- gsub("\\t", " ", line)
    if (grepl("--", line) == TRUE) {
      line <- paste(sub("--", "/*", line), "*/")
    }
    sql.string <- paste(sql.string, line)
  }
  close(con)
  return(sql.string)
}
```

## Setting up connection

```{r}
bigrquery::bq_auth() # UNCOMMENT WHEN HAVEN TOCKEN ISSUES!!!
# Establecer la conexión a BigQuery
project_id <- readline(prompt = "Enter your project ID: ")

con <- dbConnect(
  bigrquery::bigquery(),
  project = project_id
)

```
# Data Extraction

Loading queries and extracting the data.
```{r}
patients_dates <- run_query(getSQL("C:/Users/cjimenez10J/Documents/projects/r_urgencias_sepsis/r_urgencias_sepsis/Analysis/sql/patients_date.sql"))
apachescore <- run_query(getSQL("C:/Users/cjimenez10J/Documents/projects/r_urgencias_sepsis/r_urgencias_sepsis/Analysis/sql/apachescore.sql"))
sensitivitylevel <- run_query(getSQL("C:/Users/cjimenez10J/Documents/projects/r_urgencias_sepsis/r_urgencias_sepsis/Analysis/sql/sensitivitylevel.sql"))
apache_sepsis <- run_query(getSQL("C:/Users/cjimenez10J/Documents/projects/r_urgencias_sepsis/r_urgencias_sepsis/Analysis/sql/apache_sepsis.sql"))
apache_infection <- run_query(getSQL("C:/Users/cjimenez10J/Documents/projects/r_urgencias_sepsis/r_urgencias_sepsis/Analysis/sql/apache_infection.sql"))
```

# Joining 1º df
```{r}
#Unimos df
dfs_to_join <- list(patients_dates, apachescore,sensitivitylevel,apache_sepsis,apache_infection)
final_df <- Reduce(function(x,y) merge(x,y, by = "patientunitstayid", all.x = TRUE), dfs_to_join)
```

## Imputations 
```{r}
#Pasamos variable edad a númerica. Reemplazamos >89 = 90. Variable género a numérica: hombre = 0, mujer = 1
final_df <- final_df %>%
  mutate(
    age = ifelse(age == '> 89', 90, as.numeric(as.character(age))))

final_df <- final_df %>%
 mutate(
   gender = as.factor(fct_explicit_na(gender, na_level = "Unknown"))) 

final_df$gender <- relevel(final_df$gender,ref = "Male")

final_df <- final_df %>% filter(gender != "Unknown")
```

### Variable unitdischargestatus (estado del que sale de la unidad)
```{r}
library(dplyr)

final_df <- final_df %>%
  mutate(
    unitdischargestatus = ifelse(
      unitdischargestatus == 'Expired', 'Expired',
      ifelse(trimws(unitdischargestatus) == '' | unitdischargestatus == 'Alive', 'Alive', NA_character_)
    )
  )

```
### Variable hospitaladmitoffset, change name
```{r}
final_df <- final_df %>%
  mutate(hospitaladmitoffset/60) %>%
  rename(er_los_H = hospitaladmitoffset)
```
### Variable apachescore,change name and na = median
```{r}
final_df <- final_df %>%
  mutate(apachescoreIV = ifelse(is.na(apachescoreIV), median(apachescoreIV, na.rm = TRUE), apachescoreIV))
```
### Variable sensitivity:level, change na = no infection
```{r}
final_df <- final_df %>%
  mutate(
    sensitivity_level = replace_na(final_df$sensitivity_level, 'No infection')
  )
```


```{r}
#Orden de las variables
final_df_order <- final_df %>%
  select(patientunitstayid, gender,age,ethnicity,hospitalid,er_los_H,sepsis_bin,infection_bin,apacheadmissiondx,hospitaladmitsource,hospitaldischargeoffset,unittype,unitvisitnumber,unitstaytype,unitdischargeoffset,unitdischargelocation, unitdischargestatus,uniquepid,apachescoreIV,sensitivity_level)
```

### NA filling
```{r}
##Columns with na values: gender
final_df_order <- final_df_order %>% drop_na(gender) %>%
  mutate(gender = na_if(gender, '')) %>%
  filter(!is.na(gender))

```






